---
title: "Figures"
author: "Lauren Oey"
date: "4/6/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(ggthemes)
library(ggtext)

trials <- read_csv("trials.csv") %>%
  mutate(goal = paste0(goal, "estimate"),
         goal = as.factor(goal),
         cost = ifelse(cost=="unif", "linear", "quadratic"),
         cond = paste(goal, cost))

sender <- trials %>% filter(roleCurrent=="sender")
receiver <- trials %>% filter(roleCurrent=="receiver")

fontfam = "Optima"
panelcolor = "#dee4f0"
condcolors = c("#cc7667","#8b2f00","#679cd0","#003c70")
```

## general info

### condition counts

```{r}
trials %>%
  distinct(subjID, goal, cost) %>%
  count(goal, cost)
```

## correlation of sender's truth and lies

```{r}
cor.test(sender$ksay, sender$k) # r = 0.76, t(9977) = 115.21, p < 0.0001
cor(sender$ksay, sender$k)^2 # r^2 = 0.57

contrasts(sender$goal) <- contr.sum(2)
summary(lmer(ksay ~ k + (1 | subjID), data=sender))$coef
summary(lmer(ksay ~ k + goal + (1 | subjID), data=sender))$coef
```




## sender's behavior (manipulation check for difference between conditions)

```{r}
s.diff.df <- sender %>%
  mutate(kdiffs = ksay - k,
         costtype = ifelse(cost == "linear", "low", "high"),
         cost = paste(costtype, cost, "cost"),
         cost = factor(cost, levels=c("low linear cost", "high quadratic cost")),
         goal = paste(goal, "goal"))
s.bias <- sender %>%
  group_by(goal, cost) %>%
  summarise(bias.ksay = mean(ksay-k),
            median.ksay = median(ksay-k)) %>%
  mutate(bias = round(bias.ksay, 1),
         bias = ifelse(bias > 0, paste0("+",bias), as.character(bias)),
         bias = ifelse(bias == "+8", "'+8.0'", bias),
         bias = paste0("Delta[message - truth] == ", bias),
         costtype = ifelse(cost == "linear", "low", "high"),
         cost = paste(costtype, cost, "cost"),
         cost = factor(cost, levels=c("low linear cost", "high quadratic cost")),
         goal = paste(goal, "goal"))

r.diff.df <- receiver %>%
  mutate(kdiffs = kest - ksay,
         costtype = ifelse(cost == "linear", "low", "high"),
         cost = paste(costtype, cost, "cost"),
         cost = factor(cost, levels=c("low linear cost", "high quadratic cost")),
         goal = paste(goal, "goal"))
r.bias <- receiver %>%
  group_by(goal, cost) %>%
  summarise(bias.kest = mean(kest-ksay),
            median.kest = median(kest-ksay)) %>%
  mutate(bias = round(bias.kest, 1),
         bias = ifelse(bias > 0, paste0("+",bias), as.character(bias)),
         bias = paste0("Delta[estimate - message] == ", bias),
         costtype = ifelse(cost == "linear", "low", "high"),
         cost = paste(costtype, cost, "cost"),
         cost = factor(cost, levels=c("low linear cost", "high quadratic cost")),
         goal = paste(goal, "goal"))

ggplot() +
  geom_rect(data=s.diff.df, aes(xmin=-88, xmax=88, ymin=-1750, ymax=0), fill="gray40") +
  geom_vline(xintercept = 0) +
  geom_segment(data=s.bias, aes(x=bias.ksay, xend=bias.ksay, y=0, yend=1750), colour="gray60") +
  geom_segment(data=r.bias, aes(x=bias.kest, xend=bias.kest, y=0, yend=-1750), colour="gray60") +
  geom_histogram(data=s.diff.df, aes(x=kdiffs, y=..count.., fill=cond), colour="black") +
  geom_histogram(data=r.diff.df, aes(x=kdiffs, y=-..count.., fill=cond), colour="black") +
  geom_text(data=data.frame(goal="overestimate goal"), 
            aes(x=-70, y=200, label="Sender"), parse=TRUE,
            family=fontfam) +
  geom_text(data=data.frame(goal="overestimate goal"), 
            aes(x=-70, y=-200, label="Judge"), parse=TRUE, colour="white",
            family=fontfam) +
  geom_text(data=s.bias, aes(x=47, y=1000, label=bias), 
            parse=TRUE, family=fontfam) +
  geom_text(data=r.bias, aes(x=47, y=-1000, label=bias), 
            parse=TRUE, colour="white", family=fontfam) +
  scale_x_continuous("Bias", limits=c(-88,88), expand=c(0,0)) +
  scale_y_continuous("Count",
                     limits=c(-1750,1750), expand=c(0,0),
                     breaks=seq(-1500,1500,500), labels=c(seq(1500,0,-500),seq(500,1500,500))) +
  scale_fill_manual(values=condcolors) +
  facet_grid(cost ~ goal) +
  guides(fill=FALSE) +
  theme_classic() +
  theme(strip.background = element_rect(fill=panelcolor),
        strip.text = element_text(size=13, family=fontfam),
        panel.spacing = unit(0.8, "lines"),
        axis.title = element_text(size=13, family=fontfam),
        axis.text = element_text(size=11, family=fontfam))
ggsave("img/final_manuscript/player_hist.png", width=7.5, height=5)

# saving as pdf throws an "invalid font type" error
```

## check that sender lies in the right direction

```{r}
sender.stat <- sender %>%
  mutate(bias.ksay = ksay - k,
         goal = as.factor(goal),
         goal.c = goal,
         cost = as.factor(cost))

# sender overestimates > 0
sender.stat %>% # xbar = 5.59
  filter(goal == "overestimate") %>%
  summarise(mean = mean(bias.ksay))
sender.stat %>% # t(5056) = 29.09
  filter(goal == "overestimate") %>%
  pull(bias.ksay) %>%
  t.test(alternative="greater", mu=0)
# t$statistic
# t$parameter #df
# t$p.value
t.test(filter(sender.stat,goal == "overestimate")$ksay, filter(sender.stat,goal == "overestimate")$k, paired=T) #equivalent

# linear model
summary(lmer(bias.ksay ~ goal + (1|subjID), data=sender.stat))
summary(lmer(bias.ksay ~ (1|subjID) + (1|k), data=filter(sender.stat, goal=="overestimate"))) #beta = 5.98, t(162) = 5.99, p < 0.0001

# sender underestimates < 0
sender.stat %>% # xbar = -4.33
  filter(goal == "underestimate") %>%
  summarise(mean = mean(bias.ksay))
sender.stat %>% # t(4921) = -24.00
  filter(goal == "underestimate") %>%
  pull(bias.ksay) %>%
  t.test(alternative="less", mu=0)

#linear model
summary(lmer(bias.ksay ~ (1|subjID) + (1|k), data=filter(sender.stat, goal=="underestimate"))) #beta = -4.46, t(132) = -5.50, p < 0.0001

contrasts(sender.stat$goal.c) <- contr.sum(2)
# contrasts(sender.stat$cost) <- contr.treatment(2)
summary(lmer(bias.ksay ~ goal.c + (1 | subjID), data=sender.stat))
summary(lmer(bias.ksay ~ goal.c + goal:cost + (1 | subjID), data=sender.stat))
summary(lmer(bias.ksay ~ goal.c * cost + (1 | subjID), data=sender.stat))
summary(lm(bias.ksay ~ goal.c + goal:cost, data=sender.stat))

summary(lmer(ksay ~ k + goal.c + (1 | subjID), data=sender.stat))
summary(lmer(ksay ~ k + goal.c + goal:cost + (1 | subjID), data=sender.stat))
summary(lmer(ksay ~ k + goal.c * cost + (1 | subjID), data=sender.stat)) #this will do
summary(lmer(ksay ~ k + goal.c * cost + (1 | subjID), data=sender.stat))$coef

m.full <- lmer(ksay ~ k * goal.c * cost + (1 | subjID), data=sender.stat)
summary(m.full)
m.noInterKGoal <- lmer(ksay ~ k + goal.c * cost + k:cost + k:goal + (1 | subjID), data=sender.stat)
summary(m.noInterKGoal)
anova(m.full, m.noInterKGoal)
```

## check that sender lie magnitude aligns with cost functions

```{r}
s.over <- sender.stat %>%
  filter(goal == "overestimate") 
mean(filter(s.over, cost=="linear")$bias.ksay) - mean(filter(s.over, cost=="quadratic")$bias.ksay) # dbar = 5.09
t.test(filter(s.over, cost=="linear")$bias.ksay, # t(3679) = 14.13
       filter(s.over, cost=="quadratic")$bias.ksay, 
       alternative="greater", var.equal=FALSE)
contrasts(s.over$cost) <- c(1,0)
summary(lmer(bias.ksay ~ cost + (1|subjID) + (1|k), data=s.over)) #beta = 4.86, t(102) = 3.00, p = .003

s.under <- sender.stat %>%
  filter(goal == "underestimate") 
mean(filter(s.under, cost=="linear")$bias.ksay) - mean(filter(s.under, cost=="quadratic")$bias.ksay) # dbar = -5.25
t.test(filter(s.under, cost=="linear")$bias.ksay, # t(2834) = -14.16
       filter(s.under, cost=="quadratic")$bias.ksay, 
       alternative="less", var.equal=FALSE)
summary(lmer(bias.ksay ~ cost + (1|subjID) + (1|k), data=s.under)) #beta = 5.21, t(98) = 3.74, p < .001

#combined
s.comb <- s.under %>%
  mutate(bias.ksay = -bias.ksay) %>%
  bind_rows(s.over) %>%
  mutate(cost = factor(cost, levels=c("quadratic","linear")))

summary(lmer(bias.ksay ~ cost + (1|subjID) + (1|k), data=s.comb)) #beta = 5.15, t(202) = 4.81, p < 0.0001
```


## judge's considering goals (correcting in the right direction)

```{r}
receiver.stat <- receiver %>%
  mutate(bias.ksay = kest - ksay,
         goal = as.factor(goal),
         goal.c = goal,
         cost = as.factor(cost))

# receiver  kest - ksay < 0
receiver.stat %>% # xbar = -6.86
  filter(goal == "overestimate") %>%
  summarise(mean = mean(bias.ksay))
receiver.stat %>% # t(5167) = -39.85
  filter(goal == "overestimate") %>%
  pull(bias.ksay) %>%
  t.test(alternative="less", mu=0)
summary(lmer(bias.ksay ~ (1|subjID) + (1|ksay), data=filter(receiver.stat, goal=="overestimate"))) #beta = -6.84, t(148) = -7.31, p < 0.0001

contrasts(receiver.stat$goal.c) <- c(1,-1)
summary(lmer(bias.ksay ~ goal.c + (1 | subjID), data=receiver.stat))

# receiver  kest - ksay > 0
receiver.stat %>% # xbar = 5.83
  filter(goal == "underestimate") %>%
  summarise(mean = mean(bias.ksay))
receiver.stat %>% # t(3275) = 27.45
  filter(goal == "underestimate") %>%
  pull(bias.ksay) %>%
  t.test(alternative="greater", mu=0)
summary(lmer(bias.ksay ~ (1|subjID) + (1|ksay), data=filter(receiver.stat, goal=="underestimate"))) #beta = 5.33, t(159) = 7.20, p < 0.0001
```

## judge varying behavior by cost (magnitude of correcting larger for easier cost)

```{r}
r.over <- receiver.stat %>%
  filter(goal == "overestimate") 
mean(filter(r.over, cost=="linear")$bias.ksay) - mean(filter(r.over, cost=="quadratic")$bias.ksay) # dbar = -3.22
t.test(filter(r.over, cost=="linear")$bias.ksay, # t(5057) = -9.63
       filter(r.over, cost=="quadratic")$bias.ksay, 
       alternative="less", var.equal=FALSE)
summary(lmer(bias.ksay ~ cost + (1|subjID) + (1|ksay), data=r.over)) #beta = 3.21, t(102) = 2.77, p = .007

r.under <- receiver.stat %>%
  filter(goal == "underestimate") 
mean(filter(r.under, cost=="linear")$bias.ksay) - mean(filter(r.under, cost=="quadratic")$bias.ksay) # dbar = 2.50
t.test(filter(r.under, cost=="linear")$bias.ksay, # t(4080) = 7.63
       filter(r.under, cost=="quadratic")$bias.ksay, 
       alternative="greater", var.equal=FALSE)

#combined analysis
summary(lmer(bias.ksay ~ cost + (1|subjID) + (1|ksay), data=r.under)) #beta = -2.48, t(98) = -2.12, p = .037

#combined
r.comb <- r.under %>%
  mutate(bias.ksay = -bias.ksay) %>%
  bind_rows(r.over) %>%
  mutate(cost = factor(cost, levels=c("quadratic","linear")))

summary(lmer(bias.ksay ~ cost + (1|subjID) + (1|k), data=r.comb)) #beta = -2.96, t(202) = -3.53, p < 0.001
```

## no difference between sender in over vs underestimating goals
```{r}
summary(lmer(bias.ksay ~ goal + (1|subjID) + (1|ksay), 
             data=mutate(sender.stat, 
                         bias.ksay = ifelse(goal == "overestimate", -bias.ksay, bias.ksay)))) #beta = -0.67, t(200) = -1.22, p = 0.23
```

## no difference between judge in over vs underestimating goals
```{r}
summary(lmer(bias.ksay ~ goal + (1|subjID) + (1|ksay), 
             data=mutate(receiver.stat, 
                         bias.ksay = ifelse(goal == "overestimate", -bias.ksay, bias.ksay)))) #beta = -1.16, t(203) = -1.36, p = 0.18
```

## judges do not systematically overcorrect relative to sender bias
```{r}
trials.stat <- trials %>%
  mutate(bias = case_when(roleCurrent == "receiver" ~ kest - ksay,
                          roleCurrent == "sender" ~ ksay - k),
         bias.isPositive = bias > 0,
         bias = ifelse(roleCurrent == "receiver", -bias, bias),
         abs.bias = abs(bias),
         goal = as.factor(goal),
         goal.c = goal,
         cost = as.factor(cost))

summary(lmer(bias ~ roleCurrent + (1|subjID), data=trials.stat)) #no signif
summary(lmer(abs.bias ~ roleCurrent + (1|subjID), data=trials.stat)) #signif
  
full.model = lmer(abs.bias ~ goal + cost + roleCurrent + (1|subjID), data=trials.stat)
noRole.model = lmer(abs.bias ~ goal + cost + (1|subjID), data=trials.stat)
anova(full.model, noRole.model, test="LRT")
```

## reduced bias accuracy of judge estimate vs lie from heuristic.

```{r}
r.2 <- receiver %>%
  mutate(cost = paste(cost, "cost"),
         goal = paste(goal, "goal"))
abline.col = "gray40"
ai.col = "black"

ggplot(r.2, aes(x=k, y=kest, colour=cond)) +
  geom_abline(slope=1, colour=abline.col, size=0.3) +
  # geom_jitter(data=r2, aes(x=k, y=ksay), colour=ai.col, size=0.1, alpha=0.2) +
  geom_jitter(size=0.5, alpha=0.1) +
  geom_smooth(data=r.2, aes(x=k, y=ksay), method="lm", se=FALSE, size=1, colour=ai.col, alpha=0.5) +
  geom_smooth(method="lm", se=FALSE, size=1) +
  geom_hline(aes(yintercept=-Inf), colour="black", size=0.5, linetype="solid") + 
  geom_vline(aes(xintercept=-Inf), colour="black", size=0.5, linetype="solid") + 
  scale_x_continuous("Actual Truth", limits=c(0,100), expand=c(0,0)) +
  scale_y_continuous("Inferred Truth", limits=c(0,100), expand=c(0,0)) +
  scale_colour_manual(values=condcolors) +
  guides(colour=FALSE) +
  facet_grid(cost ~ goal) +
  theme_tufte() +
  theme(plot.background = element_rect(fill = "white", colour="white"),
        strip.background = element_rect(fill=panelcolor),
        strip.text = element_text(size=15, family="Optima"),
        panel.spacing = unit(1, "lines"),
        axis.title = element_text(size=15, family="Optima"),
        axis.text = element_text(size=11, family="Optima"))
ggsave("img/final/bias_lines.png", width=6,height=5.6)
```

```{r}
receiver %>%
  group_by(goal, cost) %>%
  summarise(bias.ksay = mean(ksay-k),
            bias.kest = mean(kest-k))
```

## but thats our AI heuristic... how do they compare to human, cost-sensitive liars? aggregated (mean bias) 

```{r}

```

## Comparison of judges correction and their own behavior as sender.  
does magnitude of believed bias track subjective cost for that person? (individual correlations.)
(adding more variance, maybe too confusing?)

```{r}
condcolors2 = c("#679cd0","#003c70","#cc7667","#8b2f00")
s2 <- sender %>%
  group_by(subjID, cost, goal, cond) %>%
  summarise(mean.ksay = mean(ksay-k))
c(min(s2$mean.ksay),max(s2$mean.ksay))

r2 <- receiver %>%
  group_by(subjID) %>%
  summarise(mean.kest = mean(kest-ksay))
c(min(r2$mean.kest),max(r2$mean.kest))

sr.2 <- s2 %>%
  left_join(r2) %>%
  mutate(cond = paste(cost, goal),
         cond = factor(cond,
                       levels=c("linear underestimate",
                                "quadratic underestimate",
                                "linear overestimate",
                                "quadratic overestimate")))
sr.2.summ <- sr.2 %>%
  group_by(cond) %>%
  summarise(r.mean = mean(mean.kest),
            r.se = sd(mean.kest)/sqrt(n()),
            s.mean = mean(mean.ksay),
            s.se = sd(mean.ksay)/sqrt(n()))

ggplot(sr.2, aes(x=mean.ksay, y=mean.kest, colour=cond)) +
  geom_hline(yintercept=0, colour="gray40", size=0.3) +
  geom_vline(xintercept=0, colour="gray40", size=0.3) +
  geom_point(size=1, alpha=0.3, show.legend = FALSE) +
  # geom_smooth(method="lm") +
  stat_ellipse(size=0.6) +
  geom_point(data=sr.2.summ, aes(x=s.mean, y=r.mean, fill=cond),
             shape=23, colour="black", size=2.5) +
  # geom_linerange(data=sr.2.summ, 
  #                aes(x=s.mean, y=r.mean, ymin=r.mean-r.se, ymax=r.mean+r.se), 
  #                size=0.8, show.legend = FALSE) +
  # geom_linerange(data=sr.2.summ, 
  #                aes(x=s.mean, xmin=s.mean-s.se, xmax=s.mean+s.se, y=r.mean), 
  #                size=0.8, show.legend = FALSE) +
  scale_x_continuous("Individual Sender's Mean *Report - Truth*", limits=c(-45, 45)) +
  scale_y_continuous("Individual Judge's Mean *Inferred - Report*", 
                     limits=c(-45, 45)) +
  scale_colour_manual("", values=condcolors2) +
  scale_fill_manual("", values=condcolors2) +
  # facet_wrap(~cost) +
  theme_classic() +
  theme(legend.position = c(0.74, 0.88),
        strip.background = element_rect(fill=panelcolor),
        text=element_text(size=13, family="Optima"),
        axis.title.x = element_markdown(),
        axis.title.y = element_markdown())
ggsave("img/final/indiv_lambda.png", width=5, height=4.5)
```

```{r}
# sb <- sender %>%
#   mutate(bias.ksay = ksay-k)
# rb <- receiver %>%
#   mutate(bias.kest = kest-ksay)

s3 <- sender %>%
  group_by(cost, goal) %>%
  summarise(mean.ksay = mean(ksay-k))

r3 <- receiver %>%
  group_by(cost, goal) %>%
  summarise(mean.kest = mean(kest-ksay))
comp.aggr <- s3 %>%
  left_join(r3) %>%
  rename(aggr.ksay = mean.ksay,
         aggr.kest = mean.kest)


comp.aggr
cor.test(comp.aggr$aggr.ksay, comp.aggr$aggr.kest) # r = -0.97, t(2) = -5.46, p = 0.032
```

## Individual differences

```{r}
sr.2.groups <- sr.2 %>%
  left_join(comp.aggr, by=c("cost","goal")) %>%
  mutate(corr.ksay = mean.ksay - aggr.ksay, #distance to aggregated means
         corr.kest = mean.kest - aggr.kest) 
  
  
cor.test(sr.2.groups$corr.ksay, sr.2.groups$corr.kest) # r = -0.53, t(202) = -8.84, p < 0.0001
cor(sr.2.groups$corr.ksay, sr.2.groups$corr.kest)^2

```

