---
title: "ABM"
author: "Lauren Oey"
date: "2/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(cowplot)
library(extraDistr)
library(colorspace)
library(MASS)
library(RColorBrewer)
library(scales)
library(lemon)
```

## Variables

```{r}
softmax.param = 40
# ns = seq(0,100,2)
ns = seq(0,1,0.05)
n = length(ns)
```

## General Functions

k, ksay, kest are all scaled between 0 and 1

```{r}
logitToProb <- function(logit){
  exp(logit) / (1+exp(logit))
}

probToLogit <- function(prob){
  log(prob / (1 - prob))
}

softmax <- function(alph, allEV) { ## allEV = vector of numerics
  aev = exp(alph * allEV)
  return(aev/sum(aev))
}
```



### Utility
```{r}
# sender's utility
u.S <- function(ksay, k, kest, 
  a, alph,
  m, mu){ #coop
  rw = kest - k
  mw = ksay - k
  # preserves sign when alph is even
  #rw.pow = ifelse(alph %% 2 == 0 & rw < 0 & !coop, -(rw^alph), rw^alph) 
  
  #a * rw.pow + m * abs(mw)^mu
  a * rw^alph + m * abs(mw)^mu
}


# ns2 = 0:10
# n2 = length(ns2)
# round(
#   array(
#     u.S(rep(ns2, n2*n2), rep(rep(ns2,each=n2), n2), rep(ns2,each=n2*n2),
#         1,
#         1,
#         -1,
#         0.5),
#     dim=c(n2,n2,n2)),
#   2) #x(rows) = ksay, y(cols) = k, z(lists) = kest

# receiver's utility
u.R <- function(k, kest){
  -(kest - k)^2
}

# matrix(u.R(rep(ns2, n2), rep(ns2,each=n2)), nrow=n2)
```





```{r}
## Receiver0 P(kest|ksay)
r.0 <- function() {
  diag(n)
}

## Sender P(ksay|k)
# to run with even alph, need to preserve sign (coop Boolean)
sender = function(a, alph, m, mu,
                  p.R.kest_ksay){ #x(rows) = ksay, y(cols) = k, z(list) = kest
  arr.pR <- aperm(array(rep(p.R.kest_ksay, each=n), dim=c(n,n,n)), c(3,1,2))
  arr.uS <- array(
    u.S(ns, 
        rep(rep(ns, each=n), n), 
        rep(ns, each=n*n),
        a, alph, m, mu), 
    dim=c(n,n,n))
  EV = apply(arr.uS * arr.pR, MARGIN=c(1,2), FUN=sum)
  apply(EV, MARGIN=2, softmax, softmax.param) # listsums -> 2D arr, x(rows) = ksay, y(cols) = k
}

## Receiver P(kest|ksay)
receiver = function(p.S.ksay_k){ #x(rows) = ksay, y(cols) = k, z(list) = kest
  p.S.k_ksay <- p.S.ksay_k / rowSums(p.S.ksay_k) # p(k | ksay)
  arr.pS <- array(rep(p.S.k_ksay, n), dim=c(n,n,n))
  arr.uR <- array(rep(mapply(function(i) u.R(i, ns), ns), each=n), dim=c(n,n,n))
  EV.R = t(apply(arr.uR * arr.pS, MARGIN=c(1,3), FUN=sum))
  apply(EV.R, MARGIN=2, softmax, softmax.param) # rowsums -> 2D arr, x(rows) = kest, y(cols) = ksay
}
```

kest | k

```{r}
## Receiver inference about truth P(kest|k)
receiver.kest_k = function(p.S.ksay_k, p.R.kest_ksay){
  p.S.ksay.k.arr = array(rep(p.S.ksay_k / sum(p.S.ksay_k), n), dim=c(n,n,n))
  p.R.kest_ksay.arr = aperm(array(rep(p.R.kest_ksay, each=n), dim=c(n,n,n)), c(3,1,2))
  marginalized = t(apply(p.S.ksay.k.arr * p.R.kest_ksay.arr, MARGIN=c(2,3), FUN=sum)) #marginalize over ksay
  marginalized / colSums(marginalized)
}
```

### Figure Generator
```{r}
fig <- function(mat, labx="", laby="", title="", role, leg.position="none") {
  roleColor = case_when(role=="sender" ~ "forestgreen",
                        role=="receiver" ~ "brown",
                        role=="receiver_k" ~ "sandybrown")
  mat %>%
    as_tibble() %>%
    mutate(ky = ns) %>%
    pivot_longer(-ky, names_to = 'kx', values_to='probability') %>%
    mutate(kx = (as.numeric(substr(kx, 2, (n-1)))-1) / (n-1)) %>%
    ggplot(aes(x=kx, y=ky, fill=probability)) +
    geom_tile(size=0.5) +
    geom_abline(slope=1, colour="gray70") +
    ggtitle(title) +
    scale_x_continuous(labx, expand=c(0,0), breaks=seq(0,1,0.5)) +
    scale_y_continuous(laby, expand=c(0,0), breaks=seq(0,1,0.5)) +
    scale_fill_gradient2("", 
                         low="white", mid="turquoise4", high="midnightblue", midpoint=0.5,
                         #low="white", mid="turquoise4", high="midnightblue", midpoint=0.5,
                       limits=c(0,1), labels=c("0%","25%","50%","75%","100%")) +
    theme(axis.line = element_line(colour = roleColor, size=1.5),
          panel.background = element_rect(fill = "white"),
          legend.position=leg.position)
}
#fig(agent.LN(1,1,-1,2,5)[[3]], "k", "kest", "Receiver L5 Deceptive", "receiver")

fig.horiz <- fig(p.r0, "ksay", "kest", "L0 Receiver", "receiver", "right") +
  theme(legend.direction="horizontal",
        legend.key.width=unit(2.5,"cm"))
legend_horiz <- get_legend(fig.horiz)
# 
# fig.vert <- fig(p.r0, "ksay", "kest", "L0 Receiver", "receiver", "right") +
#   theme(legend.direction="vertical",
#         legend.key.height=unit(1.5,"cm"))
# legend_vert <- get_legend(fig.vert)
# 
# ggplot(data.frame(x=c(0,1), y=c(0,1)),
#        aes(x=x, y=y)) +
#     geom_line(colour="midnightblue", size=1.5) +
#     scale_x_continuous("ksay", expand=c(0,0), breaks=seq(0,1,0.5)) +
#     scale_y_continuous("kest", expand=c(0,0), breaks=seq(0,1,0.5)) +
#     theme(axis.line = element_line(colour = "brown", size=1.5),
#           panel.background = element_rect(fill = "white"),
#           legend.position="none")
# 
# ggsave("img/l0.png", width=2, height=2)
```




# 1. Do deceptive senders and receivers reach an equilibrium? Or do they continue ratcheting to more extreme messages?

```{r}
softmax.param = 40
sim.a = 1
sim.alph = 1
sim.m = -1 #0
sim.mu = 2

p.r0 = r.0()
p.s1.defec.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r0)
p.r1.kest_ksay = receiver(p.s1.defec.ksay_k)
p.r1.kest_k = receiver.kest_k(p.s1.defec.ksay_k, p.r1.kest_ksay)
p.s2.defec.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r1.kest_ksay)
p.r2.kest_ksay = receiver(p.s2.defec.ksay_k)
p.r2.kest_k = receiver.kest_k(p.s2.defec.ksay_k, p.r2.kest_ksay)
p.s3.defec.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r2.kest_ksay)
p.r3.kest_ksay = receiver(p.s3.defec.ksay_k)
p.r3.kest_k = receiver.kest_k(p.s3.defec.ksay_k, p.r3.kest_ksay)
p.s4.defec.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r3.kest_ksay)
p.r4.kest_ksay = receiver(p.s4.defec.ksay_k)
p.r4.kest_k = receiver.kest_k(p.s4.defec.ksay_k, p.r4.kest_ksay)
p.s5.defec.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r4.kest_ksay)
p.r5.kest_ksay = receiver(p.s5.defec.ksay_k)
p.r5.kest_k = receiver.kest_k(p.s5.defec.ksay_k, p.r5.kest_ksay)

# eval.bias(p.s5.defec.ksay_k)
# eval.bias(p.r5.kest_ksay)
# eval.bias(p.r5.kest_k)
# eval.kest_k.rsq.sim(p.r5.kest_k)

blank_defec <- ggplot() + annotate("text", label="(b) Deceptive", x=1, y=1, size=6) + theme_void()
text_green <- ggplot() + annotate("text", label=bquote(atop("Sender's message given truth",P[S]~'('~k[say]~'|'~'k'~')')), x=1, y=1, size=4) + theme_void()
text_red <- ggplot() + annotate("text", label=bquote(atop("Judge's estimate given message",P[J]~'('~k[est]~'|'~k[say]~')')), x=1, y=1, size=4) + theme_void()
text_gold <- ggplot() + annotate("text", label=bquote(atop("Judge's estimate given truth",~'P'~'('~k[est]~'|'~'k'~')')), x=1, y=1, size=4) + theme_void()
fig.r0 <- fig(p.r0, "ksay", "kest", "L0 Judge", "receiver")
fig.s1 <- fig(p.s1.defec.ksay_k, "k", "ksay", "L1 Sender", "sender")
fig.r1 <- fig(p.r1.kest_ksay, "ksay", "kest", role="receiver")
fig.t1 <- fig(p.r1.kest_k, "k", "kest", role="receiver_k")
fig.s2 <- fig(p.s2.defec.ksay_k, title="L2", role="sender")
fig.r2 <- fig(p.r2.kest_ksay, role="receiver")
fig.t2 <- fig(p.r2.kest_k, role="receiver_k")
fig.s3 <- fig(p.s3.defec.ksay_k, title="L3", role="sender")
fig.r3 <- fig(p.r3.kest_ksay, role="receiver")
fig.t3 <- fig(p.r3.kest_k, role="receiver_k")
fig.s4 <- fig(p.s4.defec.ksay_k, title="L4", role="sender")
fig.r4 <- fig(p.r4.kest_ksay, role="receiver")
fig.t4 <- fig(p.r4.kest_k, role="receiver_k")
fig.s5 <- fig(p.s5.defec.ksay_k, title="L5", role="sender")
fig.r5 <- fig(p.r5.kest_ksay, role="receiver")
fig.t5 <- fig(p.r5.kest_k, role="receiver_k")

decep_levels <- plot_grid(
          text_green, text_red, text_gold,
          NULL, fig.r0, NULL,
          fig.s1, fig.r1, fig.t1, 
          fig.s2, fig.r2, fig.t2, 
          fig.s3, fig.r3, fig.t3, 
          fig.s4, fig.r4, fig.t4, 
          #fig.s5, fig.r5, fig.t5, 
          nrow=3, byrow = FALSE, rel_widths=c(2,1,1,1,1,1)) + 
  theme(plot.margin = unit(c(1,0,0,0), "cm"))
decep_levels
```


Cooperative sender and receiver inference

```{r}
sim.a = -1
sim.alph = 2
sim.m = 0
sim.mu = 0
#sim.coop = TRUE

p.r0 = r.0()
p.s1.coop.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r0)
p.r1.kest_ksay = receiver(p.s1.coop.ksay_k)
p.r1.kest_k = receiver.kest_k(p.s1.coop.ksay_k, p.r1.kest_ksay)
p.s2.coop.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r1.kest_ksay)
p.r2.kest_ksay = receiver(p.s2.coop.ksay_k)
p.r2.kest_k = receiver.kest_k(p.s2.coop.ksay_k, p.r2.kest_ksay)
p.s3.coop.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r2.kest_ksay)
p.r3.kest_ksay = receiver(p.s3.coop.ksay_k)
p.r3.kest_k = receiver.kest_k(p.s3.coop.ksay_k, p.r3.kest_ksay)
p.s4.coop.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r3.kest_ksay)
p.r4.kest_ksay = receiver(p.s4.coop.ksay_k)
p.r4.kest_k = receiver.kest_k(p.s4.coop.ksay_k, p.r4.kest_ksay)
p.s5.coop.ksay_k = sender(sim.a, sim.alph, sim.m, sim.mu, p.r4.kest_ksay)
p.r5.kest_ksay = receiver(p.s5.coop.ksay_k)
p.r5.kest_k = receiver.kest_k(p.s5.coop.ksay_k, p.r5.kest_ksay)

# eval.bias(p.s5.coop.ksay_k)
# eval.bias(p.r5.kest_ksay)
# eval.bias(p.r5.kest_k)
# eval.kest_k.rsq.sim(p.r5.kest_k)

blank_coop <- ggplot() + annotate("text", label="(a) Cooperative", x=1, y=1, size=6) + theme_void()
fig.r0 <- fig(p.r0, "ksay", "kest", "L0 Judge", "receiver")
fig.s1 <- fig(p.s1.coop.ksay_k, "k", "ksay", "L1 Sender", "sender")
fig.r1 <- fig(p.r1.kest_ksay, "ksay", "kest", role="receiver")
fig.t1 <- fig(p.r1.kest_k, "k", "kest", role="receiver_k")
fig.s2 <- fig(p.s2.coop.ksay_k, title="L2", role="sender")
fig.r2 <- fig(p.r2.kest_ksay, role="receiver")
fig.t2 <- fig(p.r2.kest_k, role="receiver_k")
fig.s3 <- fig(p.s3.coop.ksay_k, title="L3", role="sender")
fig.r3 <- fig(p.r3.kest_ksay, role="receiver")
fig.t3 <- fig(p.r3.kest_k, role="receiver_k")
fig.s4 <- fig(p.s4.coop.ksay_k, title="L4", role="sender")
fig.r4 <- fig(p.r4.kest_ksay, role="receiver")
fig.t4 <- fig(p.r4.kest_k, role="receiver_k")
fig.s5 <- fig(p.s5.coop.ksay_k, title="L5", role="sender")
fig.r5 <- fig(p.r5.kest_ksay, role="receiver")
fig.t5 <- fig(p.r5.kest_k, role="receiver_k")

coop_levels <- plot_grid(
          text_green, text_red, text_gold,
          NULL, fig.r0, NULL,
          fig.s1, fig.r1, fig.t1, 
          fig.s2, fig.r2, fig.t2, 
          fig.s3, fig.r3, fig.t3, 
          fig.s4, fig.r4, fig.t4, 
          #fig.s5, fig.r5, fig.t5, 
          nrow=3, byrow = FALSE, rel_widths=c(2,1,1,1,1,1)) + 
  theme(plot.margin = unit(c(1,0,0,0), "cm"))

plot_grid(coop_levels, decep_levels, legend_horiz, nrow=3, rel_heights=c(0.45,0.45,0.1), 
          labels=c("(a) Cooperative","(b) Deceptive"), label_fontface="plain", label_size=22,
          hjust=0)

ggsave("img/levels.png",height=11.5, width=9.5)
```



Behavior converges to equilibrium, so not using Poisson weighting...
Examine behavior of LN agent
```{r}
# depth is 0-indexed
agent.LN <- function(a, alph, m, mu, depth){
  store.s.ksay_k <- array(NA, dim=c(n,n,depth+1))
  store.r.kest_ksay <- array(NA, dim=c(n,n,depth+1))
  store.r.kest_k <- array(NA, dim=c(n,n,depth+1))
  p.r0 = r.0()
  store.s.ksay_k[,,1] = NA
  store.r.kest_ksay[,,1] = p.r0
  store.r.kest_k[,,1] = NA
  
  if(depth > 0){
    for(d in 2:(depth+1)){
      store.s.ksay_k[,,d] = sender(a, alph, m, mu, store.r.kest_ksay[,,d-1])
      store.r.kest_ksay[,,d] = receiver(store.s.ksay_k[,,d])
      store.r.kest_k[,,d] = receiver.kest_k(store.s.ksay_k[,,d], store.r.kest_ksay[,,d])
    }
  }
  
  # return last level of sender and receivers
  return(list(store.s.ksay_k[,,depth+1], # [[1]] returns depth-level sender
              store.r.kest_ksay[,,depth+1], # [[2]] returns depth-level receiver kest | ksay
              store.r.kest_k[,,depth+1])) # [[3]] returns depth-level receiver kest | k
}



#agent.LN(1,1,-1,2,3) # 3rd level deceptive agent
#agent.LN(-1,2,0,0,2) # 2nd level cooperative agent
```

# Measure variance of kest | k

```{r}
err.mat = -matrix(u.R(rep(ns, n), rep(ns,each=n)), nrow=n) #reusing receiver's L2 loss function w/ neg sign

# weighted variance of kest | k
eval.kest_k.wvar <- function(kest_k){
  sum(kest_k * err.mat) / sum(kest_k) # sum of denom should be 51
}

eval.kest_k.wvar(r.0())
eval.kest_k.wvar(agent.LN(1,1,-1,2,20)[[3]])
eval.kest_k.wvar(agent.LN(10,1,-1,2,20)[[3]])
```

# Measure R^2 of kest | k

```{r}
# eval.kest_k.rsq <- function(kest_k){
#   scols = kest_k*bias.mat - colMeans(kest_k*bias.mat)
#   srows = kest_k*bias.mat - rowMeans(kest_k*bias.mat)
#   var.kest = sum(scols^2)
#   var.k = sum(srows^2)
#   sum(scols * srows)/sum(kest_k) / sqrt(var.kest * var.k)
# }
# eval.kest_k.rsq(agent.LN(1,1,-1,2,3)[[3]])

eval.kest_k.rsq.sim <- function(kest_k){
  uncountedProb.df <- kest_k %>%
    as_tibble() %>% 
    mutate(kest = ns) %>% 
    pivot_longer(-kest, names_to = 'k', values_to='probability') %>% 
    mutate(k = (as.numeric(substr(k, 2, 10))-1) * 0.02,
           probability = round(10000 * probability)) %>%
    rename(counts = probability) %>%
    uncount(weights = counts)
  cor(uncountedProb.df$kest, uncountedProb.df$k)^2
}

eval.kest_k.rsq.sim(agent.LN(1,1,-1,2,3)[[3]])
```

# Measure bias
```{r}
direc.err <- function(k1, k2){
  k1 - k2
}
bias.mat = matrix(direc.err(rep(ns, n), rep(ns,each=n)), nrow=n)
# t(bias.mat)

eval.bias <- function(ktype){
  sum(ktype * bias.mat) / sum(ktype) # sum of denom should be 51
}

eval.bias(r.0())
eval.bias(agent.LN(1,1,-1,2,2)[[1]]) # decep sender's bias
eval.bias(agent.LN(1,1,-1,2,5)[[2]]) # decep receiver's bias
```



## Ratio of params (bias acc vs message cost) and variance kest | k



```{r}
# at ratio = 2.5 is when bias starts to converge to 0 because of extreme lies
# hold a constant at 1
ratio.combs <- data.frame(a = rep(1, each=181), 
                          ratio=c(1/seq(10,1.1,-0.1),seq(1,10,0.1))) %>% #ratio=c(1/seq(10,1.5,-0.5),seq(1,10,0.5))) %>%
  mutate(m = a*ratio)

var.decep.df = data.frame()
for(i in 1:nrow(ratio.combs)){
  a = ratio.combs$a[i]
  m = ratio.combs$m[i]
  l = 20
  agent = agent.LN(a,1,-m,2,l)
  s.ksay_k.bias = eval.bias(agent[[1]])
  r.kest_ksay.bias = eval.bias(agent[[2]])
  r.kest_k.bias = eval.bias(agent[[3]])
  r.kest_k.var = eval.kest_k.wvar(agent[[3]])
  r.kest_k.rsq = eval.kest_k.rsq.sim(agent[[3]])
  var.decep.df <- var.decep.df %>%
    bind_rows(
      data.frame(
        a = a,
        m = -m,
        l = l,
        s.ksay_k.bias = s.ksay_k.bias,
        r.kest_ksay.bias = r.kest_ksay.bias,
        r.kest_k.bias = r.kest_k.bias,
        r.kest_k.var = r.kest_k.var,
        r.kest_k.rsq = r.kest_k.rsq
      )
    )
}

# confirm min point of kest | ksay bias is at ratio = 3
var.decep.df %>%
  filter(r.kest_ksay.bias == min(r.kest_ksay.bias))

var.decep.df %>%
  filter(r.kest_k.rsq > 0.7 & r.kest_k.rsq < 0.75)

decep.df <- var.decep.df %>%
  filter(m <= -1/3) %>%
  mutate(neg.m = -m,
         logratio = log10(a/neg.m))

manual_grid_line <- data.frame(y = seq(-0.5,0.5,0.25))


decep.df_bytype <- decep.df %>%
  gather("bias_type","bias",4:6) %>%
  mutate(bias_type = case_when(bias_type == "s.ksay_k.bias" ~ "sender ksay | k", 
                               bias_type == "r.kest_ksay.bias" ~ "receiver kest | ksay",
                               bias_type == "r.kest_k.bias" ~ "receiver kest | k"),
         bias_type = factor(bias_type, levels=c("sender ksay | k","receiver kest | k","receiver kest | ksay")))

ratio.bias.fig <- decep.df_bytype %>%
  ggplot(aes(x=logratio)) +
  geom_rect(aes(xmin=-Inf, xmax=log10(3), ymin=-Inf, ymax=0), fill="gray40") +
  geom_segment(data=manual_grid_line, aes(x=-Inf, xend=Inf, y=y, yend=y), colour="gray60") +
  geom_line(aes(y=bias, colour=bias_type),size=2) +
  geom_point(aes(y=bias, colour=bias_type),size=3) +
  geom_text(data=NULL,
            aes(x=0.47, y=0.53, label="sender's bias introduced to message given the truth (ksay - k)"), 
            family="Optima", size=3.25, hjust=1) +
  geom_text(data=NULL,
            aes(x=0.47, y=-0.04, label="judge's accuracy of estimate given the truth (kest - k)"), 
            family="Optima", colour="white", size=3.25, hjust=1) +
  geom_text(data=NULL,
            aes(x=0.47, y=-0.38, label="judge's bias correction to estimate given the message (kest - ksay)"), 
            family="Optima", colour="white", size=3.25, hjust=1) +
  scale_x_continuous("Ratio of Intended Bias to Message Cost", 
                     breaks=c(-1,log10(1/3),0,log10(3)), 
                     labels=c("1 to 10", "1 to 3", "1 to 1", "3 to 1"),
                     expand=c(0,0)) +
  scale_y_continuous("Bias", limits=c(-0.53,0.53), breaks=seq(-0.5,0.5,0.25)) +
  scale_colour_manual(values=c("forestgreen","sandybrown","brown")) +
  guides(colour=FALSE) +
  theme_classic() +
  theme(plot.margin=unit(c(0.5,0.5,0.5,0.5), "cm"))

ratio.var.fig <- decep.df %>%
  ggplot(aes(x = logratio, y=r.kest_k.rsq)) +
  geom_line(size=2, colour="sandybrown") +
  geom_point(size=3, colour="sandybrown") +
  scale_x_continuous("Ratio of Intended Bias to Message Cost",
                     limits=c(-1, log10(3)), # c(-1,1),
                     breaks=c(-1,log10(1/3),0,log10(3)),
                     labels=c("1 to 10", "1 to 3", "1 to 1", "3 to 1"),
                     expand=c(0,0),
                     minor_breaks=c(log10(1/seq(10,1,-1)), log10(2))) +
  scale_y_continuous("Prop. Variance Explained (R-Squared)", breaks=seq(0,1,0.2), 
                     labels=c("   0.0", "   0.2", "   0.4", "   0.6", "   0.8", "   1.0"),
                     limits=c(0,1), expand=c(0,0)) +
  guides(colour=F) +
  theme_bw() +
  theme(plot.margin=unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.grid.minor = element_line(size=0.2))

plot_grid(ratio.bias.fig, ratio.var.fig, nrow=2, labels=c("a","b"))
ggsave("img/ratios.png",width=5,height=7.5)
```








# 2. Do cooperative senders inflate their lies when they suspect the receiver to de-bias messages?

When population is 50% deceptive, 50% cooperative?

```{r}
softmax.param = 40
p.s1.defec.ksay_k = sender(1, 1, -1, 2, p.r0)
p.s1.coop.ksay_k = sender(-1, 2, 0, 0, p.r0)
p.defec = 0.5
p.s1.comb.ksay_k = p.defec*p.s1.defec.ksay_k + (1-p.defec)*p.s1.coop.ksay_k
p.r1.kest_ksay = receiver(p.s1.comb.ksay_k)
p.s2.coop.ksay_k = sender(-1, 2, 0, 0, p.r1.kest_ksay)

fig.s1 <- fig(p.s1.comb.ksay_k, "k", "ksay", "L1 50% Deceptive Sender", role="sender")
fig.r1 <- fig(p.r1.kest_ksay, "ksay", "kest", "L2 Judge", role="receiver") 
fig.s2 <- fig(p.s2.coop.ksay_k, "k", "ksay", "L2 Cooperative Sender", role="sender")
comb50 <- plot_grid(
  plot_grid(fig.s1, fig.r1, fig.s2, ncol=1), 
  legend_vert,
  rel_widths=c(0.75,0.25))
#ggsave(paste0("img/defec",p.defec*100,".png"), width=8, height=5.5)
```

```{r}
pDefec.df <- data.frame()

p.r0 = r.0()
p.s1.defec.ksay_k = sender(1, 1, -1, 2, p.r0)
p.s1.coop.ksay_k = sender(-1, 2, 0, 0, p.r0)
for(p.defec in seq(0,1,0.005)){
  p.s1.comb.ksay_k = p.defec*p.s1.defec.ksay_k + (1-p.defec)*p.s1.coop.ksay_k
  p.r1.kest_ksay = receiver(p.s1.comb.ksay_k)
  p.s2.coop.ksay_k = sender(-1, 2, 0, 0, p.r1.kest_ksay)
  p.s2.defec.ksay_k = sender(1, 1, -1, 2, p.r1.kest_ksay)
  pDefec.df <- pDefec.df %>%
    bind_rows(
      data.frame(
        pDefec = p.defec,
        biasS2 = eval.bias(p.s2.coop.ksay_k),
        biasD2 = eval.bias(p.s2.defec.ksay_k)
      )
  )
}


pDefec.defec.df <- data.frame(pDefec = 1,
                              biasS2 = eval.bias(agent.LN(1,1,-1,2,2)[[1]]),
                              goal = "defec",
                              label = "Deceptive Sender")
round(filter(pDefec.df,pDefec==1)$biasS2,3)
fig.pDefecSumm <- ggplot(pDefec.df, aes(x=pDefec, y=biasS2)) +
  geom_point() +
  geom_line(size=2.5) +
  geom_point(data=pDefec.defec.df,
             aes(x=pDefec,y=biasS2),
             colour="red", size=3.5, shape=18) +
  geom_text(data=pDefec.defec.df,
            aes(label=label),
            hjust=1, vjust=1, nudge_x=-0.01, colour="red") +
  scale_x_continuous("Percentage of Deceptive Senders in Population",
                     limits=c(0,1), expand=c(0,0,0,0.015), labels=c("0%","25%","50%","75%","100%")) +
  scale_y_continuous("Bias of L2 Cooperative Sender", 
                     limits=c(0,0.5), expand=c(0,0)) +
  scale_colour_manual(values=c("black","red")) +
  guides(colour=FALSE) +
  coord_capped_cart(bottom='right') +
  theme_classic() +
  theme(axis.title=element_text(size=14),
        axis.text=element_text(size=12),
        plot.margin = unit(c(0.8,0.5,0,0.5), "cm"))

plot_grid(comb50, fig.pDefecSumm, nrow=1, rel_widths=c(0.3,0.7), labels=c("a","b"))
ggsave("img/defecSumm.png", width=11, height=7.5)
```
