---
title: "General Sender ABM"
author: "Lauren Oey"
date: "6/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(cowplot)
library(extraDistr)
library(colorspace)
```

## Variables

```{r}
alph = 40
ns = seq(0,1,0.02)
n = length(ns)
alph.bet = 3
```

## General Functions

k, ksay, kest are all scaled between 0 and 1

```{r}
logitToProb <- function(logit){
  exp(logit) / (1+exp(logit))
}

probToLogit <- function(prob){
  log(prob / (1 - prob))
}

softmax <- function(alph, allEV) { ## allEV = vector of numerics
  aev = exp(alph * allEV)
  return(aev/sum(aev))
}
```

### Figure Generator
```{r}
fig <- function(mat, labx="", laby="", title="") {
  mat %>%
    as_tibble() %>%
    mutate(ky = ns) %>%
    pivot_longer(-ky, names_to = 'kx', values_to='probability') %>%
    mutate(kx = (as.numeric(substr(kx, 2, (n-1)))-1) / (n-1)) %>%
    ggplot(aes(x=kx, y=ky, fill=probability)) +
    geom_tile(size=0.5) +
    ggtitle(title) +
    scale_x_continuous(labx, expand=c(0,0)) +
    scale_y_continuous(laby, expand=c(0,0)) +
    # scale_fill_gradient(low="white", high="blue")
    scale_fill_gradient2(low="white", mid="turquoise4", high="midnightblue", midpoint=0.5,
                       limits=c(0,1), labels=c("0%","25%","50%","75%","100%"))
}
```

### Priors
```{r}
p.k <- function(prior = "u"){
  if(prior == "u") {
    matrix(rep(1/n, each=n*n), nrow=n)
  } else if(prior == "b") {
    matrix(rep(dbinom(ns*10,n-1,0.5), each=n), nrow=n)
  } else if(prior == "bb") {
    matrix(rep(dbbinom(round(ns*10), n-1, alph.bet, alph.bet), each=n), nrow=n)
  }
}
```

### Utility
```{r}
# sender's utility
u.S <- function(ksay, k, kest,
  bias.rw, error.rw,
  bias.mw, error.mw,
  bias.rm, error.rm){
  rw = kest - k
  mw = ksay - k
  rm = kest - ksay
  bias.rw * rw - error.rw * rw^2 + bias.mw * mw - error.mw * mw^2 + bias.rm * rm - error.rm * rm^2
}


# round(
#   array(
#     u.S(rep(seq(0,1,0.1), 11*11), rep(rep(seq(0,1,0.1),each=11), 11), rep(seq(0,1,0.1),each=11*11), 1, 1), 
#     dim=c(n,n,n)),
#   2) #x(rows) = ksay, y(cols) = k, z(lists) = kest

# receiver's utility
u.R <- function(k, kest){
  -(kest - k)^2
}

# t(matrix(u.R(rep(seq(0,1,0.1), 11), rep(seq(0,1,0.1),each=11)), nrow=11))
```





```{r}
## Receiver0 P(kest|ksay)
r.0 <- function() {
  diag(n)
}

## Sender P(ksay|k)
sender = function(bias.rw, error.rw,
                  bias.mw, error.mw,
                  bias.rm, error.rm,
                  p.R.kest_ksay){ #x(rows) = ksay, y(cols) = k, z(list) = kest
  arr.pR <- aperm(array(rep(p.R.kest_ksay, each=n), dim=c(n,n,n)), c(3,1,2))
  arr.uS <- array(
    u.S(ns,
        rep(rep(ns, each=n), n),
        rep(ns, each=n*n),
        bias.rw, error.rw, bias.mw, error.mw, bias.rm, error.rm),
    dim=c(n,n,n))
  EV = apply(arr.uS * arr.pR, MARGIN=c(1,2), FUN=sum)
  apply(EV, MARGIN=2, softmax, alph) # listsums -> 2D arr, x(rows) = ksay, y(cols) = k
}

## Receiver P(kest|ksay)
receiver = function(p.S.ksay_k){ #x(rows) = ksay, y(cols) = k, z(list) = kest
  p.S.k_ksay <- p.S.ksay_k / rowSums(p.S.ksay_k) # p(k | ksay)
  arr.pS <- array(rep(p.S.k_ksay, n), dim=c(n,n,n))
  arr.uR <- array(rep(mapply(function(i) u.R(i, ns), ns), each=n), dim=c(n,n,n))
  EV.R = t(apply(arr.uR * arr.pS, MARGIN=c(1,3), FUN=sum))
  apply(EV.R, MARGIN=2, softmax, alph) # rowsums -> 2D arr, x(rows) = kest, y(cols) = ksay
}
```

kest | k

```{r}
## Receiver inference about truth P(kest|k)
receiver.k = function(p.S.ksay_k){ #x(rows) = ksay, y(cols) = k, z(list) = kest
  p.S.k_ksay <- p.S.ksay_k / rowSums(p.S.ksay_k) # p(k | ksay)
  arr.pS <- array(rep(p.S.k_ksay, n), dim=c(n,n,n))
  arr.uR <- array(rep(mapply(function(i) u.R(i, ns), ns), each=n), dim=c(n,n,n))
  EV.R = t(apply(arr.uR * arr.pS, MARGIN=c(2,3), FUN=sum)) # marginalizes over ksay
  apply(EV.R, MARGIN=2, softmax, alph) # rowsums -> 2D arr, x(rows) = kest, y(cols) = k
}
```


```{r}
ns1 = seq(0,0.4,0.1) #sums to 1
n1 = length(ns1)
# receiver's behavior
# x(rows) = kest
# y(cols) = ksay
# validation: cols sum to 1
(test.r.0 = array(rep(ns1, n1), dim=c(n1,n1)))
# ------------->
# x(rows) = ksay
# y(cols) = k
# z(list) = kest
# validation: test.r.0 should be repeated across cols
(test.arr.pR = aperm(array(test.r.0, dim=c(n1,n1,n1)), c(3,1,2)))

# sender's utility
# x(rows) = ksay
# y(cols) = k
# z(list) = kest
test.b.rw = 1
test.e.rw = 0
test.b.mw = 0
test.e.mw = 1
test.b.rm = 0
test.e.rm = 0
# validation: cooperative sender (0,1,0,0,0,0): cols == list ? value = 0
# validation: deceptive sender (1,0,0,1,0,0)
(test.arr.uS <- array(
    u.S(ns1, 
        rep(rep(ns1, each=n1), n1), 
        rep(ns1, each=n1*n1),
        test.b.rw, test.e.rw, test.b.mw, test.e.mw, test.b.rm, test.e.rm), 
    dim=c(n1,n1,n1)))
# sums lists (kest)
(test.EV = apply(test.arr.uS * test.arr.pR, MARGIN=c(1,2), FUN=sum))
# apply softmax on columns
(test.p.S.ksay_k <- apply(test.EV, MARGIN=2, softmax, alph))

#validation: cols sum to 1
colSums(test.p.S.ksay_k)
```

Verifying receiver function

```{r}
# sender's behavior
# x(rows) = ksay
# y(cols) = k
# validation: cols sum to 1
test.p.S.ksay_k

# p(k | ksay)
(test.p.S.k_ksay <- test.p.S.ksay_k / rowSums(test.p.S.ksay_k)) 

# ------------->
# x(rows) = ksay
# y(cols) = k
# z(list) = kest
# validation: test.p.S.k_ksay should be repeated across lists
(test.arr.pS <- array(rep(test.p.S.k_ksay, n1), dim=c(n1,n1,n1)))
# receiver always wants accuracy
(test.arr.uR <- array(rep(mapply(function(i) u.R(i, ns1), ns1), each=n1), dim=c(n1,n1,n1)))

# sums columns (k)
# then transposes
(test.EV.R = t(apply(test.arr.uR * test.arr.pS, MARGIN=c(1,3), FUN=sum)))
# apply softmax on columns
(test.p.R.kest_ksay = apply(test.EV.R, MARGIN=2, softmax, alph))

#validation: cols sum to 1
colSums(test.p.R.kest_ksay)
```

# 1. Do deceptive senders and receivers reach an equilibrium? Or do they continue ratcheting to more extreme messages?

```{r warning=FALSE}
p.r0 = r.0()
cost.mult = 1
p.s1.defec.ksay_k = sender(1, 0, 0, cost.mult, p.r0)
p.r1.kest_ksay = receiver(p.s1.defec.ksay_k)
p.s2.defec.ksay_k = sender(1, 0, 0, cost.mult, p.r1.kest_ksay)
p.r2.kest_ksay = receiver(p.s2.defec.ksay_k)
p.s3.defec.ksay_k = sender(1, 0, 0, cost.mult, p.r2.kest_ksay)
p.r3.kest_ksay = receiver(p.s3.defec.ksay_k)
p.s4.defec.ksay_k = sender(1, 0, 0, cost.mult, p.r3.kest_ksay)
p.r4.kest_ksay = receiver(p.s4.defec.ksay_k)
p.s5.defec.ksay_k = sender(1, 0, 0, cost.mult, p.r4.kest_ksay)
p.r5.kest_ksay = receiver(p.s5.defec.ksay_k)
p.s6.defec.ksay_k = sender(1, 0, 0, cost.mult, p.r5.kest_ksay)


blank <- ggplot() + theme_minimal()
fig.r0 <- fig(p.r0, "ksay", "kest", "Receiver L0") 
legend <- get_legend(fig.r0)
fig.s1 <- fig(p.s1.defec.ksay_k, "k", "ksay", "Sender L1 Deceptive") + theme(legend.position="none")
fig.r1 <- fig(p.r1.kest_ksay, "ksay", "kest", "Receiver L1") + theme(legend.position="none")
fig.s2 <- fig(p.s2.defec.ksay_k, "k", "ksay", "Sender L2 Deceptive") + theme(legend.position="none")
fig.r2 <- fig(p.r2.kest_ksay, "ksay", "kest", "Receiver L2") + theme(legend.position="none")
fig.s3 <- fig(p.s3.defec.ksay_k, "k", "ksay", "Sender L3 Deceptive") + theme(legend.position="none")
fig.r3 <- fig(p.r3.kest_ksay, "ksay", "kest", "Receiver L3") + theme(legend.position="none")
fig.s4 <- fig(p.s4.defec.ksay_k, "k", "ksay", "Sender L4 Deceptive") + theme(legend.position="none")
fig.r4 <- fig(p.r4.kest_ksay, "ksay", "kest", "Receiver L4") + theme(legend.position="none")
fig.s5 <- fig(p.s5.defec.ksay_k, "k", "ksay", "Sender L5 Deceptive") + theme(legend.position="none")
fig.r5 <- fig(p.r5.kest_ksay, "ksay", "kest", "Receiver L5") + theme(legend.position="none")
fig.s6 <- fig(p.s6.defec.ksay_k, "k", "ksay", "Sender L6 Deceptive") + theme(legend.position="none")
plot_grid(blank,fig.r0 + theme(legend.position="none"), 
          fig.s1, fig.r1, fig.s2, fig.r2, fig.s3, fig.r3, fig.s4, fig.r4, fig.s5, fig.r5, fig.s6, legend,
          nrow=2, byrow = FALSE)
ggsave("img/model.full_defec.png", width=13, height=4)
```

Receiver kest | k

```{r}
p.r0 = r.0()
cost.mult = 1
p.s1.defec.ksay_k = sender(1, 0, 0, cost.mult, 0, 0, p.r0)
p.r1.kest_ksay = receiver(p.s1.defec.ksay_k)
p.r1.kest_k = receiver.k(p.s1.defec.ksay_k)
p.s2.defec.ksay_k = sender(1, 0, 0, cost.mult, 0, 0, p.r1.kest_ksay)
p.r2.kest_ksay = receiver(p.s2.defec.ksay_k)
p.r2.kest_k = receiver.k(p.s2.defec.ksay_k)
p.s3.defec.ksay_k = sender(1, 0, 0, cost.mult, 0, 0, p.r2.kest_ksay)

blank <- ggplot() + theme_minimal()
fig.r0 <- fig(p.r0, "ksay", "kest", "Receiver L0") 
legend <- get_legend(fig.r0)
fig.s1 <- fig(p.s1.defec.ksay_k, "k", "ksay", "Sender L1 Deceptive") + theme(legend.position="none")
fig.r1 <- fig(p.r1.kest_k, "k", "kest", "Receiver L1") + theme(legend.position="none")
fig.s2 <- fig(p.s2.defec.ksay_k, "k", "ksay", "Sender L2 Deceptive") + theme(legend.position="none")
fig.r2 <- fig(p.r2.kest_k, "k", "kest", "Receiver L2") + theme(legend.position="none")
fig.s3 <- fig(p.s3.defec.ksay_k, "k", "ksay", "Sender L3 Deceptive") + theme(legend.position="none")

plot_grid(blank,fig.r0 + theme(legend.position="none"), 
          fig.s1, fig.r1, fig.s2, fig.r2, fig.s3, legend,
          nrow=2, byrow = FALSE)
```


Behavior converges to equilibrium, so not using Poisson weighting...
Examine behavior of L2 sender

Examine behavior of L1 receiver



Verifying sender function

## Ratio of params (bias acc vs message cost) and variance kest | k









# 2. Do cooperative senders inflate their lies when they suspect the receiver to de-bias messages?
When population is 100% deceptive?

```{r warning=FALSE}
p.defec = 1
p.r0 = r.0()
p.s1.defec.ksay_k = sender(1, 0, 0, 1, 0, 0, p.r0)
p.s1.coop.ksay_k = sender(0, 1, 0, 0, 0, 0, p.r0)
p.s1.comb.ksay_k = p.defec*p.s1.defec.ksay_k + (1-p.defec)*p.s1.coop.ksay_k
p.r1.kest_ksay = receiver(p.s1.comb.ksay_k)
p.s2.coop.ksay_k = sender(0, 1, 0, 0, 0, 0, p.r1.kest_ksay)


blank <- ggplot() + theme_minimal()
fig.r0 <- fig(p.r0, "ksay", "kest", "Receiver L0") 
legend <- get_legend(fig.r0)
fig.s1 <- fig(p.s1.comb.ksay_k, "k", "ksay", "Sender L1 Combined") + theme(legend.position="none")
fig.r1 <- fig(p.r1.kest_ksay, "ksay", "kest", "Receiver L1") + theme(legend.position="none")
fig.s2 <- fig(p.s2.coop.ksay_k, "k", "ksay", "Sender L2 Cooperative") + theme(legend.position="none")
plot_grid(blank,fig.r0 + theme(legend.position="none"), 
          fig.s1, fig.r1, fig.s2, legend,
          nrow=2, byrow = FALSE)
ggsave("img/model.full_coop_defec100.png", width=8, height=5.5)
```

When population is 50% deceptive, 50% cooperative?

```{r warning=FALSE}
p.defec = 0.5
p.r0 = r.0()
p.s1.defec.ksay_k = sender(1, 0, 0, 1, 0, 0, p.r0)
p.s1.coop.ksay_k = sender(0, 1, 0, 0, 0, 0, p.r0)
p.s1.comb.ksay_k = p.defec*p.s1.defec.ksay_k + (1-p.defec)*p.s1.coop.ksay_k
p.r1.kest_ksay = receiver(p.s1.comb.ksay_k)
p.s2.coop.ksay_k = sender(0, 1, 0, 0, 0, 0, p.r1.kest_ksay)


blank <- ggplot() + theme_minimal()
fig.r0 <- fig(p.r0, "ksay", "kest", "Receiver L0") 
legend <- get_legend(fig.r0)
fig.s1 <- fig(p.s1.comb.ksay_k, "k", "ksay", "Sender L1 Combined") + theme(legend.position="none")
fig.r1 <- fig(p.r1.kest_ksay, "ksay", "kest", "Receiver L1") + theme(legend.position="none")
fig.s2 <- fig(p.s2.coop.ksay_k, "k", "ksay", "Sender L2 Cooperative") + theme(legend.position="none")
plot_grid(blank,fig.r0 + theme(legend.position="none"), 
          fig.s1, fig.r1, fig.s2, legend,
          nrow=2, byrow = FALSE)
ggsave("img/model.full_coop_defec50.png", width=8, height=5.5)
```

```{r}
for(p.def in c(0, 0.5, 1)){
  p.defec = p.def
  
  p.r0 = r.0()
  p.s1.defec.ksay_k = sender(1, 0, 0, 1, 0, 0, p.r0)
  p.s1.coop.ksay_k = sender(0, 1, 0, 0, 0, 0, p.r0)
  p.s1.comb.ksay_k = p.defec*p.s1.defec.ksay_k + (1-p.defec)*p.s1.coop.ksay_k
  p.r1.kest_ksay = receiver(p.s1.comb.ksay_k)
  p.s2.coop.ksay_k = sender(0, 1, 0, 0, 0, 0, p.r1.kest_ksay)
  
  blank <- ggplot() + theme_minimal()
  fig.r0 <- fig(p.r0, "ksay", "kest", "Receiver L0") 
  legend <- get_legend(fig.r0)
  fig.s1 <- fig(p.s1.comb.ksay_k, "k", "ksay", paste0("Sender L1 ",(1-p.def)*100,"% Cooperative")) + theme(legend.position="none")
  fig.r1 <- fig(p.r1.kest_ksay, "ksay", "kest", "Receiver L1") + theme(legend.position="none")
  fig.s2 <- fig(p.s2.coop.ksay_k, "k", "ksay", "Sender L2 Cooperative") + theme(legend.position="none")
  plot_grid(blank,fig.r0 + theme(legend.position="none"), 
            fig.s1, fig.r1, fig.s2, legend,
            nrow=2, byrow = FALSE)
  ggsave(paste0("img/model.full_p", p.def*100, "_",
                "rwbmwe_vs_rwe__rwe", 
                ".png"), 
         width=9.5, height=6)
}
```